This page and the next page should be removed

### 2021-01-26 analysis/manuscript.Rmd
add new appendix

header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
---

@@

@@

How does our perception of an object change as soon as we discover what function it serves? In this study, we presented participants with unfamiliar objects that were preceded either by matching information about their function, leading to semantically informed perception, or by non-matching information, leading to naive perception. We expected semantically informed perception to differ from naive perception within the object processing hierarchy and we capitalized on the high temporal resoluation of event-related potentials (ERPs) to find out at which stage (or stages) this was indeed the case. It allowed us to show for the first time that knowledge about the function of a previously unfamiliar object alters its early cortical processing in the same trial as it is being discovered and within less than 200 ms after the presentation of the object.

@@

Previous studies have focused on these and other ERP components to investigate which processing stages are indeed influenced by knowledge about visual objects [@abdelrahman2008; @gratton2009; @maier2014; @maier2019; @rossion2002; @rossion2004; @tanaka2001; @samaha2018; @weller2019]. Together, they suggest that not only higher-level ERPs like the semantic N400 component are influenced by acquiring information about objects, but also earlier components that typically reflect bottom-up visual processing. This was the case, for instance, when participants were presented with a range of unfamiliar objects, receiving in-depth verbal descriptions about their function for half of the objects and irrelevant verbal information (i.e., cooking recipes) for the other half of the objects [@abdelrahman2008, Experiment 1]. After this learning phase, the same objects were presented in three different ERP tasks which did not require explicit access to any of the learned semantic information. In all three tasks, ERP amplitudes in response to objects for which in-depth knowledge had been acquired differed from those in response to objects for which this had not been the case. Crucially, these differences occurred not just in the N400 component, indexing a modulation of post-perceptual semantic processing, but also in the P1 component, suggesting a top-down modulation of lower-level perceptual processing. Interestingly, the ERPs in the in-depth knowledge condition even were qualitatively similar to those for untrained but well-known objects. The modulation of the P1 component has recently been replicated when the same objects were presented under circumstances of limited attentional resources in an attentional blink paradigm [@weller2019]. In this study, the (neurophysiological) differences in P1 amplitudes between objects with in-depth versus minimal knowledge were correlated with (behavioral) differences in their detection rate during the attentional blink. This can be taken as further evidence for an influence of semantic knowledge on the early visual perception of unfamiliar objects.

@@

Stimuli for Experiments 1 and 2 consisted of 240 grayscale photographs of real-world objects, 120 of which were well-known everyday objects (e.g., a bicycle, a toothbrush), serving as filler stimuli of no interest, whereas the other 120 were rare objects presumed to be unfamiliar to the majority of participants (e.g., a galvanometer, an udu drum). A list of these unfamiliar objects can be found in the appendix. All stimuli were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz. At a standardized viewing distance of 90 cm, the images of the objects subtended approximately 3.9 degrees of participants' horizontal and vertical visual angle.

@@

# Define function to compute single-trial mean ERP ampltiudes from epochs
def compute_erps(epochs, name, tmin, tmax, roi):
    # Average across electrodes in the ROI
    roi_idx = mne.pick_channels(epochs.ch_names, roi)
    epochs_roi = mne.channels.combine_channels(epochs, dict(roi=roi_idx))
    # Average across samples in the time window
    import numpy as np
    tmin_s, tmax_s = tmin / 1000, tmax / 1000
    times_idx = np.where(np.logical_and(epochs.times >= tmin_s, epochs.times <= tmax_s))
    erps = epochs_roi.get_data()[:, 0, times_idx].mean(axis=(1, 2))
    return erps

@@

# Preprocess the EEG for Epxeriment 1 in Python
map2(
  .x = list.files("data/exp1/EEG", pattern = ".vhdr", full.names = TRUE),
  .y = behav_exp1,
  ~ do.call(py$preproc, c(list(vhdr_fname = .x, metadata = .y), preproc_params))
) %>%
  # Concatenate epochs of all participants
  py$mne$epochs$concatenate_epochs() -> epochs_exp1

@@

```{r exp1-models, eval=TRUE}
# Load preprocessed single-trial data
read_tsv("analysis/export/exp1-data.tsv", col_types = cols()) %>%
  # Factorize columns
  mutate(
    part = factor(part, levels = c("I", "II", "III")),
    condition = factor(condition, levels = c(
      "Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known"
    )),
    participant = factor(participant),
    item = factor(item)
  ) %>%
  # Remove any trials excluded from conditions
  filter(condition %in% c("Informed", "Naive")) %>%
  droplevels() -> data_exp1

@@

# Trial structure
plot_trials <- function(){
  suppressMessages(require(magick))
  color_informed <- viridisLite::viridis(1, begin = 0.1)
  color_naive <- viridisLite::viridis(1, begin = 0.5)
  ggplot() +
    coord_cartesian(xlim = c(-100, 100), ylim = c(0, 100), expand = FALSE) +
    ## PART II ##
    # Title
    annotate("text", x = 0, y = 97.2, label = "Insight part", size = 14 / .pt, family = "Helvetica", fontface = "bold") +
    # Arrow
    annotate("segment", x = -30, xend = 24, y = 79, yend = 85, arrow = arrow(length = unit(0.1, "inches"))) +
    # Squares
    annotate("rect", xmin = -37, xmax = -17, ymin = 55, ymax = 75, color = "black", fill = "white") +
    annotate("rect", xmin = -19, xmax = 1, ymin = 45, ymax = 65, color = color_informed, fill = "white") +
    annotate("rect", xmin = -19, xmax = 1, ymin = 69, ymax = 89, color = color_naive, fill = "white") +
    annotate("rect", xmin = -1, xmax = 19, ymin = 59, ymax = 79, color = "black", fill = "white") +
    annotate("rect", xmin = 17, xmax = 37, ymin = 61, ymax = 81, color = "black", fill = "white") +
    annotate("text", x = -27, y = 65.3, label = "+", size = 20 / .pt, family = "Helvetica") +
    annotate("text", x = -9, y = 55, label = "potatoes\nmashing", size = 10 / .pt, family = "Helvetica", color = color_informed) +
    annotate("text", x = -9, y = 79, label = "message\nsignalling", size = 10 / .pt, family = "Helvetica", color = color_naive) +
    annotate("text", x = 9, y = 67, label = "*", size = 30 / .pt, family = "Helvetica") +
    draw_image("analysis/manuscript_files/potato_masher.png", x = 18, y = 62, width = 18, height = 18) +
    # Timings
    annotate("text", x = -9, y = 67.3, label = "or", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = -27, y = 52, label = "0.5 s", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = -9, y = 42, label = "2.5 s", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = 9, y = 56, label = "0.5 s", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = 27, y = 56.3, label = "3 s or\nresponse", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    # Response options
    annotate("segment", x = 27, y = 39, xend = 27, yend = 51.5) +
    annotate("segment", x = 0, y = 39, xend = 27, yend = 39) +
    annotate("segment", x = 0, y = 35, xend = 0, yend = 39) +
    annotate("segment", x = -32, y = 35, xend = 32, yend = 35) +
    annotate("segment", x = -32, y = 30.5, xend = -32, yend = 35) +
    annotate("segment", x = -12, y = 32, xend = -12, yend = 35) +
    annotate("segment", x = 12, y = 32, xend = 12, yend = 35) +
    annotate("segment", x = 32, y = 32, xend = 32, yend = 35) +
    annotate("text", x = -32, y = 26, label = "A. \"I know\nwhat this is\"", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    annotate("text", x = -12, y = 26, label = "B. \"I have an\nassumption\nwhat this is\"", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    annotate("text", x = 12, y = 26, label = "C. \"I have rather\nno assumption\nwhat this is\"", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    annotate("text", x = 32, y = 26, label = "D. \"I don\'t\n know what\nthis is\"", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    # Conditions
    annotate("segment", x = -32, y = 17, xend = -32, yend = 22, color = color_informed) +
    annotate("segment", x = -12, y = 17, xend = -12, yend = 20, color = color_informed) +
    annotate("segment", x = 12, y = 17, xend = 12, yend = 20, color = color_naive) +
    annotate("segment", x = 32, y = 17, xend = 32, yend = 20, color = color_naive) +
    annotate("segment", x = -32, y = 17, xend = -12, yend = 17, color = color_informed) +
    annotate("segment", x = 12, y = 17, xend = 32, yend = 17, color = color_naive) +
    annotate("segment", x = -22, y = 15, xend = -22, yend = 17, color = color_informed) +
    annotate("segment", x = 22, y = 15, xend = 22, yend = 17, color = color_naive) +
    annotate("text", x = -22, y = 13, label = "Informed condition", size = 10 / .pt, family = "Helvetica", fontface = "bold", color = color_informed) +
    annotate("text", x = 22, y = 13, label = "Naive condition", size = 10 / .pt, family = "Helvetica", fontface = "bold", color = color_naive) +
    annotate("text", x = -22, y = 7.9, label = "Matching keywords\nand response A or B", size = 10 / .pt, family = "Helvetica", color = color_informed, lineheight = 1) +
    annotate("text", x = 22, y = 7.9, label = "Non-matching keywords\nand response C or D", size = 10 / .pt, family = "Helvetica", color = color_naive, lineheight = 1) +
    ## PART I ##
    # Title
    annotate("text", x = -72, y = 56, label = "Pre-insight part", size = 14 / .pt, family = "Helvetica", fontface = "bold") +
    # Arrow
    annotate("segment", x = -84, xend = -66, y = 46, yend = 48, arrow = arrow(length = unit(0.1, "inches"))) +
    # Squares
    annotate("rect", xmin = -91, xmax = -71, ymin = 22, ymax = 42, color = "black", fill = "white") +
    annotate("rect", xmin = -53, xmax = -73, ymin = 24, ymax = 44, color = "black", fill = "white") +
    annotate("text", x = -81, y = 32.3, label = "+", size = 20 / .pt, family = "Helvetica") +
    draw_image("analysis/manuscript_files/potato_masher.png", x = -72, y = 25, width = 18, height = 18) +
    # Timings
    annotate("text", x = -81, y = 19, label = "0.5 s", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = -63, y = 19.3, label = "3 s or\nresponse", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    # Conditions
    annotate("text", x = -72, y = 7.9, label = "Objects classified according\nto conditions in the insight part", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    # PART III
    # Title
    annotate("text", x = 72, y = 56, label = "Post-insight part", size = 14 / .pt, family = "Helvetica", fontface = "bold") +
    # Arrow
    annotate("segment", x = 60, xend = 78, y = 46, yend = 48, arrow = arrow(length = unit(0.1, "inches"))) +
    # Squares
    annotate("rect", xmin = 53, xmax = 73, ymin = 22, ymax = 42, color = "black", fill = "white") +
    annotate("rect", xmin = 71, xmax = 91, ymin = 24, ymax = 44, color = "black", fill = "white") +
    annotate("text", x = 63, y = 32.3, label = "+", size = 20 / .pt, family = "Helvetica") +
    draw_image("analysis/manuscript_files/potato_masher.png", x = 72, y = 25, width = 18, height = 18) +
    # Timings
    annotate("text", x = 63, y = 19, label = "0.5 s", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = 81, y = 19.3, label = "3 s or\nresponse", size = 10 / .pt, family = "Helvetica", lineheight = 1) +    
    # Conditions
    annotate("text", x = 72, y = 7.9, label = "Objects classified according\nto conditions in the insight part", size = 10 / .pt, family = "Helvetica", lineheight = 1) +    
    theme_void()
}

# Define function to plot ERP waveforms and topographies
plot_erps <- function(comps, evokeds, models, montage) {
  suppressWarnings(suppressMessages(
    pmap(comps, function(name, tmin, tmax, roi) {
      # N400 gets a different scale than P1 and N170
      ymin <- ifelse(name == "N400", -3, -4)
      # Loop through parts
      parts <- map(c("I", "II", "III"), function(part) {
        # Create (conditions x time points) x electrodes tibble
        data_plot <- evokeds[[part]] %>%
          map(function(data) {
            data %>%
              t() %>%
              as_tibble(.name_repair = "unique") %>%
              set_colnames(montage$electrode) %>%
              mutate(.time = seq(-500, 1498, 2), roi = rowMeans(select(., all_of(roi))))
          }) %>% bind_rows(.id = "condition")
        # Shade background depending on whether the effect is significant or not
        asterisks <- models[[name]]$contrasts %>%
          filter(part == !!part) %>%
          mutate(asterisks = case_when(p.value < .001 ~ "***", p.value < .01 ~ "**", p.value < .05 ~ "*")) %>%
          pull(asterisks)
        xasterisks <- case_when(str_length(asterisks) == 1 ~ tmin + 11, str_length(asterisks) == 3 ~ tmin + 15)
        if (is.na(asterisks)) {
          shade <- annotate("rect", xmin = tmin, xmax = tmax, ymin = ymin + 0.1, ymax = ymin + 11.95, color = "black", fill = NA)
        } else {
          shade <- annotate("rect", xmin = tmin, xmax = tmax, ymin = ymin + 0.1, ymax = ymin + 11.95, color = "black", fill = "gray90")
        }

        # # Prepare waveform
        # wave <- ggplot(data = data_plot, aes(x = .time, y = roi, color = condition)) +
        #   annotate("rect", xmin = tmin, xmax = tmax, ymin = -Inf, ymax = ymin + 12, fill = "gray90")
        # # Shade significant area
        # star <- stars[[name]][[part]]
        # if (!is.na(star)) {
        #   wave <- wave + geom_ribbon(
        #     data = data_plot %>% filter(.time >= tmin & .time <= tmax) %>%
        #       group_by(.time) %>% summarize(ymin = min(roi), ymax = max(roi)),
        #     aes(x = .time, xmin = tmin, xmax = tmax, ymin = ymin, ymax = ymax),
        #     inherit.aes = FALSE, fill = viridisLite::viridis(n = 1, begin = 1)
        #   )
        # } +
        # annotate("text", label = star, x = tmin + (tmax - tmin)/2, y = -3.2, size = 7, family = "Helvetica")}
        # Add waves and styling

        # Plot waveform
        wave <- ggplot(data = data_plot, aes(x = .time, y = roi, color = condition)) +
          shade +
          annotate("text", label = asterisks, x = xasterisks, y = ymin + 11, size = 6, hjust = 0, family = "Helvetica") +
          annotate("segment", x = -200, xend = 800, y = 0, yend = 0) +
          annotate("segment", x = 0, xend = 0, y = ymin, yend = ymin + 12) +
          annotate("segment", x = seq(-100, 700, 200), xend = seq(-100, 700, 200), y = -0.3, yend = 0) +
          annotate("segment", x = -12, xend = 0, y = seq(-2, 8, 4), yend = seq(-2, 8, 4)) +
          annotate("text", x = seq(-100, 700, 200), y = -0.9, label = seq(-100, 700, 200), size = 10 / .pt, family = "Helvetica") +
          annotate("text", x = -20, y = seq(-2, 8, 4), label = seq(-2, 8, 4), size = 10 / .pt, family = "Helvetica", hjust = 1) +
          annotate("text", x = 400, y = ymin + 0.8, label = "Time (ms)", size = 10 / .pt, family = "Helvetica", lineheight = 0.9) +
          annotate("text", x = -100, y = 4, label = paste(name, "ampl.\n(µV)"), size = 10 / .pt, family = "Helvetica", angle = 90, lineheight = 0.9) +
          geom_line() +
          scale_color_viridis_d(begin = 0.1, end = 0.5) +
          coord_cartesian(xlim = c(-200, 800), ylim = c(ymin, ymin + 14), expand = FALSE) +
          theme_void() +
          theme(legend.position = "none")
        # Plot topography
        topo <- data_plot %>%
          filter(.time >= tmin & .time <= tmax) %>%
          group_by(condition) %>%
          summarise(across(montage$electrode, mean)) %>%
          pivot_longer(-condition) %>%
          pivot_wider(names_from = condition) %>%
          transmute(amplitude = Informed - Naive) %>%
          bind_cols(montage) %>%
          eegUtils::topoplot(limits = c(-1, 1), palette = "viridis", contour = FALSE, highlights = roi, scaling = 0.1) +
          theme(legend.position = "none")
        topo$layers[[3]]$aes_params$size <- topo$layers[[4]]$aes_params$size <- topo$layers[[5]]$aes_params$size <- 0.6
        # Combine waveform and topography
        wave + draw_plot(topo, width = 320, height = 9, x = 500, y = ymin + 6.4)
      })
      # Combine plots for the different parts
      plot_grid(parts[[1]], NULL, parts[[2]], NULL, parts[[3]], nrow = 1, rel_widths = c(10, 1, 10, 1, 10))
      # Combine plots for the different components
    }) %>% plot_grid(plotlist = ., nrow = 3)
  ))
}

# Create a combined legend and colorbar
plot_legends_erps <- function(direction) {
  if (direction == "vertical") {
    mar <- margin(l = 12, r = 10, b = 6, t = 7); vjust <- 3.5
  }
  else {
    mar <- margin(l = 8, r = 12, b = 4, t = 6); vjust <- 0.7
  }
  get_legend(
    tibble(amplitude = c(-1, 0, 1), condition = c("Informed", "Naive", "Informed")) %>%
      ggplot(aes(x = 0, y = 0, color = condition, fill = amplitude)) +
      geom_line() +
      geom_raster() +
      labs(color = "Conditions", fill = "Informed - naive (µV)") +
      scale_color_viridis_d(begin = 0.1, end = 0.5, guide = guide_legend(direction = "vertical", title.position = "top")) +
      scale_fill_viridis_c(guide = guide_colorbar(direction = "horizontal", ticks = FALSE, title.position = "top", title.vjust = vjust)) +
      theme(
        legend.direction = direction,
        legend.box = direction,
        legend.box.background = element_rect(colour = "black", size = 0.5),
        legend.box.margin = mar,
        legend.background = element_blank(),
        legend.title = element_text(family = "Helvetica", size = 10, face = "bold"),
        legend.text = element_text(family = "Helvetica", size = 10),
        legend.key = element_rect(fill = NA),
        legend.key.width = unit(0.7, "cm"),
        legend.key.height = unit(0.4, "cm"),
        legend.spacing.x = unit(0.3, "cm")
      )
  )
}

@@

The event-related potentials in response to objects from both conditions and all three parts were analyzed on the single trial level using linear mixed-effects regression models [@baayen2008; @frömer2018]. For the purpose of the present study, these models have at least three desirable properties compared to traditional approaches such as analyses of variance (ANOVAs) performed on by-participant grand averages. First, they can leverage all of the information that is present in the data set on the single trial level, some of which would be lost when computing averages. Second, they can account simultaneously for the non-independence of data points coming from the same participant and/or from the same item. In contrast, the neglect of the item as a random variable in by-participant ANOVAs leads to anti-conservative test statistics and strictly does not allow for inferences beyond the stimulus set under study [@bürki2018; @judd2012]. Finally, mixed-effects models can flexibly deal with unbalanced designs in which the number of trials differs across (combinations of) conditions. Such a situation is inevitable when the assignment of trials to conditions is co-determined by the experimental manipulation and the behavioral responses of the participants to this manipulation [e.g., @fröber2017].

Three separate models were computed predicting P1, N170, and N400 mean amplitudes, respectively. All models included three fixed effects: (a) the part of the experiment, coded as a repeated contrast (i.e., subtracting the first from the second part and the second from the third part, the intercept being the grand mean across all three parts), (b) the condition of the object, coded as a scaled sum contrast (i.e., subtracting the naive condition from the semantically informed condition, the intercept being the grand mean across both conditions), and (c) the two-way interaction of part and condition. For details on these and other contrast coding schemes in linear (mixed-effects) models, please refer to @schad2020. To determine the random effects structure, we always started with a maximal model containing by-participant and by-item random intercepts and random slopes for all fixed effects [@barr2013]. We then performed a model selection algorithm as proposed by @matuschek2017 in order to increase statistical power and avoid overparameterization: Iteratively, each random effect was removed and the resulting, more parsimonious model was compared to the previous, more complex model by means of a likelihood ratio test. Only if the parsimonious model explained the data equally well as the complex model [determined by *p* \> .20\; @matuschek2017] did we leave the random effect out, otherwise it was kept in the final model. All models were computed in R [Version `r as.character(packageVersion("base"))`\; @R-base] using the lme4 package [Version `r as.character(packageVersion("lme4"))`\; @R-lme4]. The optimizer function *bobyqa* with 20,000 iterations was used for maximum likelihood estimation. The model selection algorithm via likelihood ratio tests was performed using the buildmer package [Version `r as.character(packageVersion("buildmer"))`\; @R-buildmer].

@@

The materials, single trial behavioral and ERP data, and all code for data analysis can be accessed via the Open Science Framework (<https://osf.io/uksbc/>). Participants' lack of informed consent to do so prevents us from openly sharing the raw EEG data.

@@

```{r exp1-table, include=TRUE, results="asis"}
# Define function to print ANOVA-style table
create_table <- function(models, stub_anova, stub_contrasts, caption, note) {
  anov <- map(models, function(model) {
    data.frame(
      "f" = format(round(model$anova$`F value`, 2), trim = FALSE, nsmall = 2),
      "df" = paste0("(", model$anova$NumDF, ", ", format(round(model$anova$DenDF, 1), trim = TRUE, nsmall = 1), ")"),
      "p" = format(round(model$anova$`Pr(>F)`, 3), trim = FALSE, nsmall = 3)
    ) %>%
      mutate(p = ifelse(p == "0.000", "< .001", substr(p, 2, nchar(p)))) %>%
      set_rownames(c(stub_anova))
  })
  suppressMessages(
    anov_print <- anov %>%
      map(unite, col = fdf, f, df, sep = " ") %>%
      map(add_row, fdf = "\\textit{F} (\\textit{df})", p = "\\textit{p}", .before = 1) %>%
      bind_cols() %>%
      set_rownames(c("\\textbf{Fixed effects}", stub_anova))
  )
  suppressMessages(
    anov %<>%
      bind_cols() %>%
      set_colnames(paste(rep(names(models), each = length(names(models))), c("f", "df", "p"), sep = "_"))
  )
  conts <- map(models, function(model) {
    data.frame(
      "est" = format(round(model$contrasts$estimate, 2), trim = FALSE, nsmall = 2),
      "ci" = paste0(
        "[", format(round(model$contrasts$lower.CL, 2), trim = TRUE, nsmall = 2), ", ",
        format(round(model$contrasts$upper.CL, 2), trim = TRUE, nsmall = 2), "]"
      ),
      "p" = format(round(model$contrasts$`p.value`, 3), trim = FALSE, nsmall = 3)
    ) %>%
      mutate(p = ifelse(p == "0.000", "< .001", substr(p, 2, nchar(p)))) %>%
      set_rownames(stub_contrasts)
  })
  suppressMessages(
    conts_print <- conts %>%
      map(unite, col = estci, est, ci, sep = " ") %>%
      map(add_row, estci = "Est. [95% CI]", p = "\\textit{p}", .before = 1) %>%
      bind_cols() %>%
      set_rownames(c("\\textbf{Informed - naive}", stub_contrasts))
  )
  suppressMessages(
    conts %<>%
      bind_cols() %>%
      set_colnames(paste(rep(names(models), each = length(names(models))), c("est", "ci", "p"), sep = "_"))
  )
  cnames <- c("\\textbf{Fixed effects}", as.character(anov_print[1, ]))
  list(anov_print[2:nrow(anov_print), ], conts_print) %>%
    map(set_colnames, paste0("V", 1:ncol(anov_print))) %>%
    apa_table(
      col.names = cnames, col_spanners = list("\\textbf{P1}" = 2:3, "\\textbf{N170}" = 4:5, "\\textbf{N400}" = 6:7),
      midrules = nrow(anov_print), font_size = "footnotesize", align = "lcccccc", escape = FALSE,
      caption = caption, note = note
    ) %>%
    cat()
  return(list("anov" = anov, "conts" = conts))
}

# Create table of models for Experiment 1
table_exp1 <- create_table(
  models = models_exp1,
  stub_anova = c("Part", "Condition", "Pt. × con."),
  stub_contrasts = c("Pre-insight part", "Insight part", "Post-insight part"),
  caption = "Results of Linear Mixed-Effects Regression Models for Experiment 1\\smallskip",
  note = "Pt. = part, con. = condition, est. = estimate, CI = confidence interval."
)
```

>>>>

header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \usepackage{makecell}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
---

@@

# Source plotting functions
source("analysis/manuscript_files/functions_plots_tables.R")

@@

Does our perception of an object change as soon as we discover what function it serves? In this study, we presented participants with unfamiliar objects that were preceded either by matching information about their function, leading to semantically informed perception, or by non-matching information, leading to naive perception. We expected semantically informed perception to differ from naive perception within the object processing hierarchy and we capitalized on the high temporal resolution of event-related potentials (ERPs) to find out at which stage (or stages) this was indeed the case. It allowed us to show for the first time that knowledge about the function of a previously unfamiliar object alters its early cortical processing in the same trial as it is being discovered and within less than 200 ms after the presentation of the object.

@@

Previous studies have focused on these and other ERP components to investigate which processing stages are indeed influenced by knowledge about visual objects [@abdelrahman2008; @gratton2009; @maier2014; @maier2019; @rossion2002; @rossion2004; @tanaka2001; @samaha2018; @weller2019]. Together, they suggest that not only higher-level ERPs like the semantic N400 component are influenced by acquiring information about objects, but also earlier components that typically reflect bottom-up visual processing. This was the case, for instance, when participants were presented with a range of unfamiliar objects, receiving in-depth verbal descriptions about their function for half of the objects and irrelevant verbal information (i.e., cooking recipes) for the other half of the objects [@abdelrahman2008, Experiment 1]. After this learning phase, the same objects were presented in three different ERP tasks which did not require explicit access to any of the learned semantic information. In all three tasks, ERP amplitudes in response to objects for which in-depth knowledge had been acquired differed from those in response to objects for which this had not been the case. These differences occurred not just in the N400 component, indexing a modulation of post-perceptual semantic processing, but also in the P1 component, suggesting a top-down modulation of lower-level perceptual processing. Interestingly, the ERPs observed for previously unfamiliar objects with newly acquired in-depth knowledge were qualitatively similar to those for untrained but well-known objects. The modulation of the P1 component has recently been replicated when the same objects were presented under circumstances of limited attentional resources in an attentional blink paradigm [@weller2019]. In this study, the (neurophysiological) differences in P1 amplitudes between objects with in-depth versus minimal knowledge were correlated with (behavioral) differences in their detection rate during the attentional blink. This can be taken as further evidence for an influence of semantic knowledge on the early visual perception of unfamiliar objects.

@@

Stimuli for Experiments 1 and 2 consisted of 240 grayscale photographs of real-world objects, 120 of which were well-known everyday objects (e.g., a bicycle, a toothbrush), serving as filler stimuli of no interest, whereas the other 120 were rare objects presumed to be unfamiliar to the majority of participants (e.g., a galvanometer, an udu drum). A list of these unfamiliar objects can be found in Appendix A. All stimuli were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz. At a standardized viewing distance of 90 cm, the images of the objects subtended approximately 3.9 degrees of participants' horizontal and vertical visual angle.

@@

# Define function to compute single-trial mean ERP amplitudes from epochs
def compute_erps(epochs, name, tmin, tmax, roi):
    # Average across electrodes in the ROI
    roi_idx = mne.pick_channels(epochs.ch_names, roi)
    epochs_roi = mne.channels.combine_channels(epochs, dict(roi=roi_idx))
    # Average across samples in the time window
    import numpy as np
    tmin_s, tmax_s = tmin / 1000, tmax / 1000
    times_idx = np.where(np.logical_and(epochs.times >= tmin_s, epochs.times <= tmax_s))
    erps = epochs_roi.get_data()[:, 0, times_idx].mean(axis=(1, 2))
    return erps

@@

# Preprocess the EEG for Experiment 1 in Python
map2(
  .x = list.files("data/exp1/EEG", pattern = ".vhdr", full.names = TRUE),
  .y = behav_exp1,
  ~ do.call(py$preproc, c(list(vhdr_fname = .x, metadata = .y), preproc_params))
) %>%
  # Concatenate epochs of all participants
  py$mne$epochs$concatenate_epochs() -> epochs_exp1

@@

```{r exp1-models, eval=rerun_models}
# Load preprocessed single-trial data
read_tsv("analysis/export/exp1-data.tsv", col_types = cols()) %>%
  # Factorize columns
  mutate(
    part = factor(part, levels = c("I", "II", "III")),
    condition = factor(condition, levels = c(
      "Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known"
    )),
    participant = factor(participant),
    item = factor(item)
  ) %>%
  # Remove any trials excluded from conditions
  filter(condition %in% c("Informed", "Naive")) %>%
  droplevels() -> data_exp1

@@

@@

The event-related potentials in response to objects from both conditions and all three parts were analyzed on the single trial level using linear mixed-effects regression models [@baayen2008; @frömer2018]. For the purpose of the present study, these models have at least three desirable properties compared to traditional approaches such as analyses of variance (ANOVAs) performed on by-participant averages. First, they can leverage all of the information that is present in the data set on the single trial level, some of which would be lost when computing averages. Second, they can account simultaneously for the non-independence of data points coming from the same participant and/or from the same item. In contrast, the neglect of the item as a random variable in by-participant ANOVAs leads to anti-conservative test statistics and strictly does not allow for inferences beyond the stimulus set under study [@bürki2018; @judd2012]. Finally, mixed-effects models can flexibly deal with unbalanced designs in which the number of trials differs across (combinations of) conditions. Such a situation is inevitable when the assignment of trials to conditions is co-determined by the experimental manipulation and the behavior of participants in response to this manipulation [e.g., @fröber2017].

Three separate models were computed predicting P1, N170, and N400 mean amplitudes, respectively. All models included three fixed effects: (a) the part of the experiment, coded as a repeated contrast (i.e., subtracting the first from the second part and the second from the third part, the intercept being the grand mean across all three parts), (b) the condition of the object, coded as a scaled sum contrast (i.e., subtracting the naive condition from the semantically informed condition, the intercept being the grand mean across both conditions), and (c) the two-way interaction of part and condition. For details on these and other contrast coding schemes in linear (mixed-effects) models, please refer to @schad2020. To determine the random effects structure, we always started with a maximal model containing by-participant and by-item random intercepts and random slopes for all fixed effects [@barr2013]. We then performed a model selection algorithm as proposed by @matuschek2017 in order to increase statistical power and avoid overparameterization: Iteratively, each random effect was removed and the resulting, more parsimonious model was compared to the previous, more complex model by means of a likelihood ratio test. Only if the parsimonious model explained the data equally well as the complex model [determined by *p* \> .20\; @matuschek2017] did we leave the random effect out, otherwise it was kept in the final model (see Appendix B for the formulas after model selection). All models were computed in R [Version `r as.character(packageVersion("base"))`\; @R-base] using the lme4 package [Version `r as.character(packageVersion("lme4"))`\; @R-lme4]. The optimizer function *bobyqa* with 20,000 iterations was used for maximum likelihood estimation. The model selection algorithm via likelihood ratio tests was performed using the buildmer package [Version `r as.character(packageVersion("buildmer"))`\; @R-buildmer].

@@

The materials, single trial behavioral and ERP data, and all code for data analysis can be accessed via the Open Science Framework (<https://osf.io/uksbc/>). We are not able to share the raw EEG data openly because we did not obtain participants' informed consent to do so.

@@

```{r exp1-table, include=TRUE, results="asis"}
# Create table of models for Experiment 1
table_exp1 <- create_table(
  models = models_exp1,
  stub_anova = c("Part", "Condition", "Pt. × con."),
  stub_contrasts = c("Pre-insight part", "Insight part", "Post-insight part"),
  caption = "Results of Linear Mixed-Effects Regression Models for Experiment 1\\smallskip",
  note = "Pt. = part, con. = condition, est. = estimate, CI = 95\\% confidence interval."
)
```

Averaged across conditions, P1, N170, and N400 amplitudes differed as a function of the part of the experiment, all *F*s ≥ `r table_exp1$anov["Part", "P1_f"]`, all *p*s `r table_exp1$anov["Part", "P1_p"]` (see Table \@ref(tab:exp1-table)). In addition, N400 amplitudes differed between the informed and the naive condition averaged across the three parts of the experiment, *F*`r table_exp1$anov["Condition", "N400_df"]` = `r table_exp1$anov["Condition", "N400_f"]`, *p* = `r table_exp1$anov["Condition", "N400_p"]`. Crucially, the part × condition interaction was significant in the N170 component, *F*`r table_exp1$anov["Pt. × con.", "N170_df"]` = `r table_exp1$anov["Pt. × con.", "N170_f"]`, *p* = `r table_exp1$anov["Pt. × con.", "N170_p"]`, and in the N400 component, *F*`r table_exp1$anov["Pt. × con.", "N400_df"]` = `r table_exp1$anov["Pt. × con.", "N400_f"]`, *p* `r table_exp1$anov["Pt. × con.", "N400_p"]`, while also showing a weak statistical trend in the P1 component, *F*`r table_exp1$anov["Pt. × con.", "P1_df"]` = `r table_exp1$anov["Pt. × con.", "P1_f"]`, *p* = `r table_exp1$anov["Pt. × con.", "P1_p"]`. To answer our main research question, we decomposed these interactions into the differences between the semantically informed condition and the naive condition within the three different parts of the experiment.
